---
title: "DESeq2 - Mesenchymal vs Embryonal"
output:
  pdf_document: default
  html_notebook: default
---

HOW TO GET FOLD CHANGE DATA USING DESeq2

Load the libraries:

```{r}
library(DESeq2)
library(readr)
```

Convert the .csv files to matrices within R

1) "countdata" (hb1): Read counts, [m x n] with m = samples, n = genes
2) "coldata" (colname): Sample assigments, [l x m] with l = variables (eg. patient, histology, treatment)

Set up the colname matrix 

```{r}
setwd ("/Users/pengwu/Documents/Research - Old/RNASeqData/180805")
hb1 <- read_csv("180805_countdata_minusoutliers.csv")
hb1_colnames <- read_csv("180805_colnames_minusoutliers.csv")
data.frame (hb1_colnames)
colname = data.frame (hb1_colnames)
```

Define the experimental design

```{r}
dds <- DESeqDataSetFromMatrix(countData = hb1,
                              colData = colname,
                              design = ~ Type)

#Collapse 'technical' replicates
ddsCollapsed <- collapseReplicates( dds,
                                    groupby = dds$Sample)

# Running the pipeline to obtain log2 fold change for mes/emb:

dds <- DESeq(ddsCollapsed)
res <- results( dds, contrast = c("Type", "Mesenchymal", "Embryonal") )
res
```

Convert to csv file (with genes numbered rather than gene names), where gene names can be added back from the original read.csv file:

```{r}
setwd ("/Users/pengwu/Documents/Experiments/RNASeq/180805Final")
write.csv(as.data.frame(res),
          file="MesvEmb_All_minusoutliersreps_Type.csv")
```


Plotting the histogram of p-values:

```{r}
hist( res$pvalue, breaks=20, col="grey" )
```

```{r}
hist( res$padj, breaks=20, col="grey" )
```
